Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-06-18 manuscript_results.Rmd
code review line 1800

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales") 
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
library("lme4")
library("ez")
```

@@

Datasets with differences and Wilcoxon Holm adjusted p-values

@@

# View the updated dataframe
head(p_values_and_impr)
p_values_and_impr
```

@@

new_long_data <- summary_impr %>%
  mutate(
    LogBack = grepl("_LB$", Model),
    Family = sub("_.*", "", Model),  # take everything before the first underscore
    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
  )

new_long_data <- new_long_data %>%
  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))

model <- lm(m ~ Model, data = new_long_data)
summary(model)

model
new_long_data$Model
# Get tidy summary of coefficients
tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
tidy_model_no_intercept <- tidy_model_no_intercept %>%
  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
                    ymax = estimate + std.error * 1.96),
                width = 0.2) +
  coord_flip() +  # flip for better readability
  labs(
    title = "Effect Size of Model Types on outcome",
    subtitle =  "Estimates relative to baseline (intercept)",
    y = "Estimate",
    x = ""
  ) +
  theme_minimal()

@@

@@

new_long_data <- summary_impr %>%
  mutate(
    LogBack = grepl("_LB$", Model),
    Type = sub("_.*", "", Model),  # take everything before the first underscore
    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
  )

new_long_data <- new_long_data %>%
  mutate(Family = ifelse(Family == "AR", "AAR", Family))

model <- lm(m ~ Type + Family + LogBack, data = new_long_data)
summary(model)

model
new_long_data$Model
# Get tidy summary of coefficients
tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
tidy_model_no_intercept <- tidy_model_no_intercept %>%
  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
                    ymax = estimate + std.error * 1.96),
                width = 0.2) +
  coord_flip() +  # flip for better readability
  labs(
    title = "Effect size of model type, family and transformation on WIS difference",
    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
    y = "WIS Difference",
    x = ""
  ) +
  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))

@@

library(ggplot2)
library(dplyr)
library(forcats)

>>>>

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("scales") 
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
library("lme4")
library("ez")
```

@@

Datasets with mean differences, standard deviations and Wilcoxon Holm adjusted p-values

@@

# View the updated dataframe
head(p_values_and_impr)
p_values_and_impr
```
ERASE

@@

#new_long_data <- summary_impr %>%
#  mutate(
#    LogBack = grepl("_LB$", Model),
#    Family = sub("_.*", "", Model),  # take everything before the first underscore
#    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
#  )

#new_long_data <- new_long_data %>%
#  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))

#model <- lm(m ~ Model, data = new_long_data)
#summary(model)

#model
#new_long_data$Model
# Get tidy summary of coefficients
#tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
#tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
#tidy_model_no_intercept <- tidy_model_no_intercept %>%
#  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
#ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
#  geom_point() +
#  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
#                    ymax = estimate + std.error * 1.96),
#                width = 0.2) +
#  coord_flip() +  # flip for better readability
#  labs(
#    title = "Effect Size of Model Types on outcome",
#    subtitle =  "Estimates relative to baseline (intercept)",
#    y = "Estimate",
#    x = ""
#  ) +
#  theme_minimal()

@@

Now let's use a regression model to evaluate effect size of each modelling approach.

@@

model_approaches <- summary_impr %>%
  mutate(
    LogBack = grepl("_LB$", Model),
    Type = sub("_.*", "", Model),  # take everything before the first underscore
    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
  )

model_approaches <- model_approaches %>%
  mutate(Family = ifelse(Family == "AR", "AAR", Family)) # rename because the intercept is in alphabetic order and I want to use the baseline

# run the regression model
model <- lm(m ~ Type + Family + LogBack, data = model_approaches)

# extracting models coefficients
tidy_model <- broom::tidy(model)

# filter out intercept
tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Reordering for plotting
tidy_model_no_intercept <- tidy_model_no_intercept %>%
  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
                    ymax = estimate + std.error * 1.96),
                width = 0.2) +
  coord_flip() + 
  labs(
    title = "Effect size of model type, family and transformation on WIS difference",
    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
    y = "WIS Difference",
    x = ""
  ) +
  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))

@@
